Freenet statistics
========================================================

Welcome to my Freenet statistics page!


```{r echo=FALSE}
opts_chunk$set(comment=NA, fig.width=20, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE)
```

```{r setup, results='hide'}
library(ggplot2)
library(xtable)
library(scales)     # Need the scales package
library(reshape)

setwd("~/freenet/retention/")

fetch <- read.csv("fetch.csv",header=T, sep="\t")
insert <- read.csv("insert.csv",header=T, sep="\t")
groups <- read.csv("group.csv",header=T, sep="\t")

fetch$date <- as.POSIXct(fetch$date, format = "%Y.%m.%d")
fetch$date <- as.Date.POSIXct(fetch$date)
fetch$version <- as.factor(fetch$version)
fetch$delta <- as.factor(fetch$delta)

insert$date <- as.POSIXct(insert$date, format = "%Y.%m.%d")
insert$date <- as.Date.POSIXct(insert$date)
insert$version <- as.factor(insert$version)


groups$version <- as.factor(groups$version)
groups$mandatory <- as.factor(groups$mandatory)

data <- merge(x=fetch,y=groups, by="version", all.x=T)
data <- na.omit(data)

insert <- merge(x=insert,y=groups, by="version", all.x=T)
```

Fetch statistics
--------------------------------------------------

### Description 
These results are based on fetching and inserting 32 single blocks of data. 
Every day 32 single blocks are inserted into Freenet. At various delays we try to fetch these blocks again. More specifically, ach day we try to retrieve the 32 blocks inserted (2^delta)-1 days ago for delta = 1,2, ... 8. All the fetch results which have been obtained with a specific version of freenet are part of the same sample. From these tests we get (1) results about the retrievability of data, (2) how long it takes for requests to complete and (3) how long inserts take.

The following graphs plot the average fetch performance for a time period. Keep in mind that we cannot clearly distinguish between fetch related improvements and insertion related improvements, i.e. they are co-dependent in Freenet. In addition to these pretty graphs we'd also like to know whether a Freenet version is substantially different from another one. Freenet has two types of `versions': *regular versions* and *mandatory versions*. Mandatory versions are assumed to substantially alter the network behaviour of Freenet, whereas regular builds do not. The statistics gathered for Freenet are grouped by mandatory version, because we assume that regular versions have no network impact. For each mandatory version we test whether all the results obtained using that version are significantly different from those obtained with another version. This test is performed using a linear regression model.

The following R output may be a bit difficult to grasp, so hang on. Here we train a linear regression model on a bunch of samples. Each sample is based on one of the last 10 mandatory builds. Each build is treated as a factor. Obviously, we would like to see a build that improves on past performance significantly with a P < 0.05. For each delta (time past since the last attempt to fetch it) the output of the linear regression model is shown. This results in 8 models in total, one for each delta. Want to know whether a specific build is an improvement compared to another one? Hunt for versions with a P < 0.05.

Now, onwards to the actual results!

### Fetch graphs
```{r fetchplots, results='asis'}
plotit <- function(datai, prefix){
  for(deltaValue in seq(1:8))
  {
    cat(paste('<h4> Plot and linear regression model for data inserted ', (2^deltaValue)-1, 'days ago</h4>\n'))
    
    plot <- ggplot(subset(datai, delta==deltaValue), aes(date, ratio, group=1, fill=mandatory, color=mandatory)) +
      geom_smooth(aes(label="smoothed data"), color="grey", alpha=0.08, span=0.3, se=F) +
      geom_boxplot(aes(group=mandatory), alpha=0.4) +
      geom_jitter() + 
      ggtitle(paste("Fetch ratio for the last", length(last_mandatory), "mandatory builds for data inserted", (2^deltaValue)-1, "day(s) ago.")) + 
      ylab("Ratio of succesfully fetched blocks")
  
    print(plot)

    print(xtable(summary(lm(ratio ~ 1 + mandatory+date, data=subset(datai, delta==deltaValue)))), type="html", comment=FALSE)
    }
}

last_mandatory = tail(unique(data$mandatory), 10)
last_builds <- subset(data, mandatory %in% last_mandatory)
plotit(last_builds, "last_builds_fetch_delta_")
```

### How long does it take for a request to complete?

The following graph plots how long it takes for an average request for a single block of data to complete.

```{r request_time}

#plot request times for different versions over time for the last 10 mandatory builds
last_10_builds <- subset(data, mandatory %in% tail(unique(data$mandatory), 10))
ggplot(last_10_builds, aes(date, time, group=delta, color=delta)) + 
  geom_smooth(span=0.2, size=2, alpha=0.6, se=F) +
  geom_jitter(size=2, alpha=0.3) +
  ylab("Fetch duration in milliseconds")

```

### How long does it take to insert stuff?

```{r insert_time }
#plot request times for different versions over time for the last 10 mandatory builds
last_10_insert_builds <- subset(insert, mandatory %in% tail(unique(data$mandatory), 10))
last_10_insert_builds$time = last_10_insert_builds$time / 1000;
ggplot(last_10_insert_builds, aes(date, time, group=mandatory, color=mandatory)) + 
  geom_boxplot() +
  geom_jitter(alpha=0.3) +
  ylab("Insert duration in seconds")
```

```{r insert_time_table, results='asis'}
print(xtable(summary(lm(time ~ 1 + mandatory+date, data=last_10_insert_builds))), type="html", comment=FALSE)
```


Bootstrapping
-----------------------------

The following graph plots how long it takes for a node to bootstrap. That means, the time in seconds required to get 6 opennet peers or more via seed nodes. This data is, at present, pulled from the main Freenet developer (toad) who tries to bootstrap a node every day and checks how long this takes.


```{r bootstrapping}
bootstrapping <- read.table("bootstrapping/data.txt",
                            header=F, 
                            sep="\t")
colnames(bootstrapping) <- c("date", "time")

bootstrapping$date <- as.Date(bootstrapping$date)

xa <- ylab("Time required to get 6 seednodes in seconds")

ggplot(data=subset(bootstrapping, date > Sys.Date() - 14), aes(date, time)) +
  geom_smooth(aes(date, time), method="loess") + 
  geom_jitter(aes(date, time), alpha=0.2) + xa


ggplot(data=bootstrapping, aes(date, "bootstrapping.time")) +
  geom_smooth(aes(date, time), span=0.1, method="loess") + 
  geom_jitter(aes(date, time), alpha=0.2) + xa

```

Seed node statistics
--------------------------------------------

The following graphs are based on data gathered from my own seednode and may not be representative for the network at large.


### Opennet freenet versions connected to my seednode


```{r versions}
versions <- read.table("version_distribution/3_days.txt",
                            header=T, 
                            sep="\t",
                            check.names = F)
versions$time <- as.POSIXct(versions$time, origin = "1970-01-01")
versions <- melt(versions, id="time", variable_name = "version")

ggplot(data=versions, aes(time, value, group=version, color=version)) +
  geom_smooth(method="loess", span=0.3) + 
  geom_jitter(alpha=0.2) + 
  ylab("# peers with this version")
```

### Number of Announcements and related messages sent

```{r announcements}
announcements <- read.csv("announcements/plot_data.txt", sep="\t", header=TRUE, colClasses = c("numeric", "numeric","numeric", "numeric"))

announcements$Time <- as.POSIXct(announcements$Time, origin="1970-01-01")
announcements <- subset(announcements, Time > Sys.time() - 3600*24*14)
announcements <- melt(announcements, id="Time", variable_name = "type")


ggplot(data=announcements, aes(Time,value, color=type)) + 
  geom_smooth(method="loess", span=0.3) + 
  geom_jitter(aes(Time, value, color=type), alpha=0.2) + 
  scale_y_continuous(trans=log2_trans()) + 
  ylab("number of peers on a log2 scale")
```


### Peer status of connected opennet peers

```{r peer_status}
connected <- read.table("peer_stats/connection_history.txt",
                            header=T, 
                            sep="\t")
                            
connected$Time <- as.POSIXct(connected$Time, origin="1970-01-01")
connected <- subset(connected, Time > Sys.time() - 3600*24*14)
connected <- melt(connected, id="Time", variable_name="Connection status")

ggplot(data=connected, aes(Time,value, group=`Connection status`, color=`Connection status`)) + 
  geom_smooth(method="loess", span=0.2) + 
  geom_jitter(aes(Time, value, color=`Connection status`, group=`Connection status`), alpha=0.15) +
  ylab("Number of connected peers")
```


Author: digger3 - USK@zALLY9pbzMNicVn280HYqS2UkK0ZfX5LiTcln-cLrMU,GoLpCcShPzp3lbQSVClSzY7CH9c9HTw0qRLifBYqywY,AQACAAE/WebOfTrust/1895 .

I tend to hang around on Sone and freenode IRC 
